### 2018-06-04 Manuscript Figures.Rmd
Updating manuscript figures

---
title: "Manuscript Figures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval.after = 'fig.cap')
```

```{r importSynMapData, include=FALSE}
source("./R/do.R")

library(ggplot2)
library(tidyr)
library(dplyr)
library(VennDiagram)
library(gridExtra)
```

# 1 Chromosome location of Maize1 and Maize2

```{r chromosome_plot, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=4}
v4_subg_jw <-
  subgenome %>%
  left_join(positions, by=c("gene2"="geneID")) %>%
  select(Chromosome, Pos_start, Pos_end, subgenome) %>%
  rename(Subgenome=subgenome) %>%
  subset(Subgenome %in% c("sub1", "sub2"))

v4_dom_jw <- expressedPairs
v4_dom_jw$Subgenome <- "no dominance"
v4_dom_jw$Subgenome[v4_dom_jw$FPKM_mean1 / v4_dom_jw$FPKM_mean2 > 2] <- "dom"
v4_dom_jw$Subgenome[v4_dom_jw$FPKM_mean2 / v4_dom_jw$FPKM_mean1 > 2] <- "dom"
v4_dom_jw$Maize1[v4_dom_jw$FPKM_mean2 / v4_dom_jw$FPKM_mean1 > 2] <- v4_dom_jw$Maize2[v4_dom_jw$FPKM_mean2 / v4_dom_jw$FPKM_mean1 > 2]
df0 <- v4_dom_jw[v4_dom_jw$Subgenome %in% c("no dominance"),]
df0$Maize1 <- df0$Maize2
v4_dom_jw <- bind_rows(v4_dom_jw, df0)
v4_dom_jw <-
  v4_dom_jw %>%
  select(Maize1, Subgenome) %>%
  left_join(positions, by=c("Maize1"="geneID")) %>%
  select(Chromosome, Pos_start, Pos_end, Subgenome)

## 6,773 - 3,875 (no map) -~-  1,018 (no positions); split/merge is fine = 1,896 orphan v4_ids with positions based on Zea_mays.AGPv4.36.chr.gff3
v4_orph_jw <-
  OrphanGenes %>%
  select(gene) %>%
  left_join(maize.genes.v3_to_v4.map, c("gene" = "v3_id"))
v4_orph_na <- v4_orph_jw[is.na(v4_orph_jw$v4_id),]
v4_orph_jw <- v4_orph_jw[!is.na(v4_orph_jw$v4_id),]
v4_orph_jw <-
  v4_orph_jw %>%
  select(v4_id) %>%
  left_join(positions, by=c("v4_id"="geneID"))
v4_orph_jw <- v4_orph_jw[!is.na(v4_orph_jw$Pos_end),]
v4_orph_jw$Subgenome <- "orphan"
v4_orph_jw <-
  v4_orph_jw %>%
  select(Chromosome, Pos_start, Pos_end, Subgenome)

centromere.v4.positions <-
  centromere.v4.positions %>%
  rename(Chromosome="X__1", Pos_start="X__6", Pos_end="X__7") %>%
  select(Chromosome, Pos_start, Pos_end)
centromere.v4.positions

df1 <- centromere.v4.positions
df1$x <- -.45
df1$xend <- .45
df2 <- v4_subg_jw
df2$x <- -.45
df2$xend <- -.05
df3 <- v4_dom_jw
df3$x <- .05
df3$xend <- .45
segment_data <- bind_rows(df2,df3)
segment_data <- segment_data[!is.na(segment_data$Pos_start),]
segment_data <-
  segment_data %>%
  ungroup()
segment_data <- segment_data %>% subset(Subgenome %in% c("sub1", "sub2", "orphan", "sub1dom", "sub2dom", "dom"))
centromere_data <- df1

#--------------------------------------------------------------------------------------------------#
## Make the plot
p <-
  ggplot(data=v4) +
    geom_bar(aes(v4$chromosome, v4$size), stat="identity", fill="grey70") +
    # geom_segment(data=segment_data, aes(x=Chromosome+x, xend=Chromosome+xend, y=Pos_start, yend=Pos_end, colour=Subgenome), size=.001) +
    geom_rect(data=segment_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end, color=Subgenome, fill=Subgenome)) +
    geom_rect(data=centromere_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end)) +
    scale_fill_manual(values=c("green", "red", "blue", "black")) +
    scale_color_manual(values=c("green", "red", "blue", "black")) +
    scale_x_discrete(limits=c("Chr1","Chr2","Chr3","Chr4","Chr5","Chr6","Chr7","Chr8","Chr9","Chr10")) +
    scale_y_continuous(labels=format_si())
p  +
  labs(
  # title="Subgenome Locations by Chromosome",
  x="Chromosome",
  y="Position on Chromosome"
)

```

# 2 GO Term Comparison by Subgenome
***

GO Terms by subgenome.  If the GO Term is assigned to any gene in the subgenome, it is counted once for that subgenome.  This is a presence/absence test.

The main takeaways here are that 1) there are differences in functional assignments between each subgenome, regardless of which filters I apply.  2) the GOSlim for plants is extremely restrictive, and probably not useful.  3) Sub1 has more functional assignments than sub2, although this may be a direct correlation to the larger size (by definition) of sub1.

GO Term assignments to genes by subgenome, which takes into account the gene-GO Term pairings and considers homeologs equivalent for the purposes of determining overlap in assignments.

The main takeaways here mirror the assessment from the GO Term presence/absence check above. There are many differences in assignments, although we can assume that the same terms are reused very often based on the differenced between unique terms and gene annotations.

```{r GOTerms_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
#EXP only, MF terms only
grid.newpage()
g <- draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
  area2 = goAnnotations.sub2 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct(),  
                         goAnnotations.sub2 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct()
                         ) %>% nrow(), 
  category = c("Maize1", "Maize2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, MF only)")

## Only keep genes where there is a duplicate still on subgenome 1 and subgenome 2.  The converted data.frame will hold maize2 genes with the names replaced with maize1 homologs
converted <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  right_join(., goAnnotations.sub2,by=c("Maize2"="gene2"))
converted$Maize2[!is.na(converted$Maize1)] <- as.character(converted$Maize1[!is.na(converted$Maize1)])
converted <- 
  converted %>%
  select(Maize2, goTerm, evCode, type) %>%
  rename(gene2=Maize2)

#EXP only, MF only
grid.newpage()
g <- draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% nrow(),
  area2 = converted %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(gene2, goTerm) %>% distinct(), 
                         converted %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(gene2, goTerm) %>% distinct()
                         ) %>% nrow(), 
  category = c("Maize1", "Maize2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="GO Term Assignments by Subgenome\n(EV-EXP, MF only)")
```

# 3 Compare Homeolog Pairs

***

Here is a scatter plot of isoform counts across gene pairs.

```{r subgenome_isoforms_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.align='center', fig.cap = cap}
# Isoforms
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(geneTranscript.counts, by=c("Maize1"="gene")) %>%
  inner_join(geneTranscript.counts, by=c("Maize2"="gene"))
names(data)[3] <- "Value_maize1"
names(data)[4] <- "Value_maize2"
# data <- subset(data, data$Value_maize1 > 1 & data$Value_maize2 > 1) ## Ignore single-exon genes
data$foldChange <- log2(data$Value_maize1) - log2(data$Value_maize2)

n <-
  data %>%
  # subset(Value_maize1 > 100 | Value_maize2 > 100) %>%
  subset(Value_maize1 > 25 | Value_maize2 > 25) %>%
  nrow()

plot <-
  ggplot(subset(data, Value_maize1 <= 25 & Value_maize2 <= 25), aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(position = "jitter", alpha=.1, aes(color = abs(foldChange))) +
  scale_color_gradient(low = "blue", high = "red") +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  # xlim(0,100) +
  # ylim(0,100) +
  xlim(0,25) +
  ylim(0,25) +
  labs(
    # title = paste0("Scatterplot of isoform counts for gene pair(s).\nExcludes ",n," outlier, alpha=.1, foldchange in log2, blue=lm fit."),
    x = "Isoform Counts for Maize1 Gene",
    y = "Isoform Counts for Maize2 Gene"
  ) + 
  # annotate("text",x=85,y=20, label=paste("Pearson Cor = ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
  annotate("text",x=20,y=3, label=paste("Pearson Cor = ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
plot(plot)

cap <- "Fig 3.1: "
```

```{r subgenome_expression_scatter_rep, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=4, fig.align='center', fig.cap = cap}
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

data <- 
  data %>%
  subset(sample %in% c("Primary Root 5 Days", "Mature Leaf 8", "6-7 internode"))

plot <-
  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(alpha=.1, aes(color = sample)) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  scale_x_log10() + scale_y_log10() +
  labs(
    # title = paste0("Scatterplot of expression for gene pair(s).\nAlpha=.1, reference line"),
    x = "Expression for Maize1 Gene (log scale)",
    y = "Expression for Maize2 Gene (log scale)"
  ) +
  theme(legend.position = "none") + facet_wrap(~sample, nrow=1)
plot(plot)

cap <- "Fig 3.2: "
```

```{r subgenome_expression_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

plot <-
  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(alpha=.1, aes(color = sample)) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  scale_x_log10() + scale_y_log10() +
  labs(
    # title = paste0("Scatterplot of expression for gene pair(s).\nAlpha=.1, reference line"),
    x = "Expression for Maize1 Gene (log scale)",
    y = "Expression for Maize2 Gene (log scale)"
  ) +
  theme(legend.position = "none") + facet_wrap(~sample, nrow=6)
plot(plot)

cap <- "Fig 3.3: "
```

```{r subgenome_protein_scatter_rep, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=4, fig.align='center', fig.cap = cap}
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))
data <- 
  data %>%
  subset(sample %in% c("Primary Root 5 Days", "Mature Leaf 8", "6-7 internode"))

plot <-
  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(alpha=.1, aes(color = sample)) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  scale_x_log10() + scale_y_log10() +
  labs(
    # title = paste0("Scatterplot of protein abundance for gene pair(s).\nAlpha=.1, reference line"),
    x = "Abundance for Maize1 Gene (log scale)",
    y = "Abundance for Maize2 Gene (log scale)"
  ) +
  theme(legend.position = "none") + facet_wrap(~sample, nrow=1)
  # + annotate("text", size=2, x=10000,y=10, label=paste("pearson: ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
plot(plot)

cap <- "Fig 3.4: "
```

```{r subgenome_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

plot <-
  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(alpha=.1, aes(color = sample)) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  scale_x_log10() + scale_y_log10() +
  labs(
    # title = paste0("Scatterplot of protein abundance for gene pair(s).\nAlpha=.1, reference line"),
    x = "Abundance for Maize1 Gene (log scale)",
    y = "Abundance for Maize2 Gene (log scale)"
  ) +
  theme(legend.position = "none") + facet_wrap(~sample, nrow=6)
  # + annotate("text", size=2, x=10000,y=10, label=paste("pearson: ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
plot(plot)

cap <- "Fig 3.5: "
```

# 4 Expression and Abundance are Weakly Correlated

```{r subgenome_expression_protein_foldchange_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=6, fig.height=5, fig.align='center', fig.cap = cap}
## Start with the paired expression data
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "FPKM_maize1"
names(data)[5] <- "FPKM_maize2"

## Add paired abundance data
data <-
  data %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID", "sample"="sample")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[6] <- "dNSAF_maize1"
names(data)[7] <- "dNSAF_maize2"

data$foldChange_expr <- log2(data$FPKM_maize1) - log2(data$FPKM_maize2)
data$foldChange_abun <- log2(data$dNSAF_maize1) - log2(data$dNSAF_maize2)
ggplot(data, aes(foldChange_expr, foldChange_abun)) +
  geom_point(alpha=.1) +
  xlim(-10,10) + ylim(-10,10) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  # facet_wrap(~sample, nrow=6) +
  labs(
    # title = paste0("Scatterplot of Foldchange in Expression in Maize1 vs.\nFoldchange in Abundance in Maize1\nAlpha=.1, linear fit, n=13413"),
    x = "log2(Expression Maize1) - log2(Expression Maize2)",
    y = "log2(Abundance Maize1) - log2(Abundance Maize2)"
  ) + 
  annotate("text",x=7,y=-8, label=paste("Pearson Cor = ", round(cor(data$foldChange_expr, data$foldChange_abun, method = c("pearson")),2)))

cap <- "Fig 10.1: Considering only homeologous pairs where both genes were measured for expression and protein abundance, we look at fold change from maize2 expression to maize1 expression vs. the fold change from maize2 abundance vs. maize2 abundance. A slightly positive trend indicates a mild positive correlation between an increase in expression and an increase in abundance."
```

```{r subgenome_expression_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
## Start with the paired expression data
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "FPKM_maize1"
names(data)[5] <- "FPKM_maize2"

## Add paired abundance data
data <-
  data %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID", "sample"="sample")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[6] <- "dNSAF_maize1"
names(data)[7] <- "dNSAF_maize2"

## View as scatter
ggplot(data, aes(FPKM_maize1, dNSAF_maize1)) +
  geom_point(alpha=.1, aes(color = sample)) +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  # geom_smooth(method="loess", se=TRUE, fullrange=FALSE, level=0.95) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  scale_x_log10() + scale_y_log10() +
  theme(legend.position = "none") + facet_wrap(~sample, nrow=6) +
  labs(
    # title = paste0("Scatterplot of Expression in Maize1 vs. Abundance in Maize1\nLog-Log Scale for gene pair(s).\nAlpha=.1, reference line (black), lm fit (blue)"),
    x = "Expression for Maize1 Gene (log scale)",
    y = "Abundance for Maize1 Gene (log scale)"
  )

cap <- "Fig 10.2: "
```

# 5 Expression pattern types
functionalDivergence.R

# 6 Wrapup
***
Report generated on: `r as.character(format(Sys.Date(), format="%B %d, %Y"))`