### 2018-06-06 Manuscript Figures.Rmd
Updated figures

---
title: "Manuscript Figures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval.after = 'fig.cap')
```

```{r importSynMapData, include=FALSE}
source("./R/do.R")

library(ggplot2)
library(tidyr)
library(dplyr)
library(VennDiagram)
library(gridExtra)
```

# 1 Chromosome location of Maize1 and Maize2

***

```{r chromosome_plot, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=8, fig.height=6}
v4_subg_jw <-
  subgenome %>%
  left_join(positions, by=c("gene2"="geneID")) %>%
  select(Chromosome, Pos_start, Pos_end, subgenome) %>%
  rename(Subgenome=subgenome) %>%
  subset(Subgenome %in% c("sub1", "sub2"))

v4_dom_jw <- expressedPairs
v4_dom_jw$Subgenome <- "no dominance"
v4_dom_jw$Subgenome[v4_dom_jw$FPKM_mean1 / v4_dom_jw$FPKM_mean2 > 2] <- "dom"
v4_dom_jw$Subgenome[v4_dom_jw$FPKM_mean2 / v4_dom_jw$FPKM_mean1 > 2] <- "dom"
v4_dom_jw$Maize1[v4_dom_jw$FPKM_mean2 / v4_dom_jw$FPKM_mean1 > 2] <- v4_dom_jw$Maize2[v4_dom_jw$FPKM_mean2 / v4_dom_jw$FPKM_mean1 > 2]
df0 <- v4_dom_jw[v4_dom_jw$Subgenome %in% c("no dominance"),]
df0$Maize1 <- df0$Maize2
v4_dom_jw <- bind_rows(v4_dom_jw, df0)
v4_dom_jw <-
  v4_dom_jw %>%
  select(Maize1, Subgenome) %>%
  left_join(positions, by=c("Maize1"="geneID")) %>%
  select(Chromosome, Pos_start, Pos_end, Subgenome)

## 6,773 - 3,875 (no map) -~-  1,018 (no positions); split/merge is fine = 1,896 orphan v4_ids with positions based on Zea_mays.AGPv4.36.chr.gff3
v4_orph_jw <-
  OrphanGenes %>%
  select(gene) %>%
  left_join(maize.genes.v3_to_v4.map, c("gene" = "v3_id"))
v4_orph_na <- v4_orph_jw[is.na(v4_orph_jw$v4_id),]
v4_orph_jw <- v4_orph_jw[!is.na(v4_orph_jw$v4_id),]
v4_orph_jw <-
  v4_orph_jw %>%
  select(v4_id) %>%
  left_join(positions, by=c("v4_id"="geneID"))
v4_orph_jw <- v4_orph_jw[!is.na(v4_orph_jw$Pos_end),]
v4_orph_jw$Subgenome <- "orphan"
v4_orph_jw <-
  v4_orph_jw %>%
  select(Chromosome, Pos_start, Pos_end, Subgenome)

centromere.v4.positions <-
  centromere.v4.positions %>%
  rename(Chromosome="X__1", Pos_start="X__6", Pos_end="X__7") %>%
  select(Chromosome, Pos_start, Pos_end)
centromere.v4.positions

df1 <- centromere.v4.positions
df1$x <- -.45
df1$xend <- .45
df2 <- v4_subg_jw
df2$x <- -.45
df2$xend <- -.05
df3 <- v4_dom_jw
df3$x <- .05
df3$xend <- .45
segment_data <- bind_rows(df2,df3)
segment_data <- segment_data[!is.na(segment_data$Pos_start),]
segment_data <-
  segment_data %>%
  ungroup()
segment_data$Subgenome[segment_data$Subgenome %in% c("sub1")] <- "Maize1"
segment_data$Subgenome[segment_data$Subgenome %in% c("sub2")] <- "Maize2"
# segment_data$Subgenome[segment_data$Subgenome %in% c("dom")] <- "Dominant Homeolog"
segment_data <- segment_data %>% subset(Subgenome %in% c("Maize1", "Maize2", "orphan", "sub1dom", "sub2dom", "dom"))
centromere_data <- df1

#--------------------------------------------------------------------------------------------------#
## Make the plot
chromosome_plot <-
  ggplot(data=v4) +
    geom_bar(aes(v4$chromosome, v4$size), stat="identity", fill="grey70") +
    # geom_segment(data=segment_data, aes(x=Chromosome+x, xend=Chromosome+xend, y=Pos_start, yend=Pos_end, colour=Subgenome), size=.001) +
    geom_rect(data=segment_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end, color=Subgenome, fill=Subgenome)) +
    geom_rect(data=centromere_data, inherit.aes = F, aes(xmin=Chromosome+x, xmax=Chromosome+xend, ymin=Pos_start, ymax=Pos_end)) +
    scale_fill_manual(values=c("green", "red", "blue", "black")) +
    scale_color_manual(values=c("green", "red", "blue", "black")) +
    scale_x_discrete(limits=c("Chr1","Chr2","Chr3","Chr4","Chr5","Chr6","Chr7","Chr8","Chr9","Chr10")) +
    scale_y_continuous(labels=format_si())
chromosome_plot  +
  labs(
  # title="Subgenome Locations by Chromosome",
  x="Chromosome",
  y="Position on Chromosome"
)

```

# 2 GO Term Comparison by Subgenome

***

```{r GOTerms_by_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4}
#EXP only, MF terms only
grid.newpage()
GOTerms_by_subgenome_a <- draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
  area2 = goAnnotations.sub2 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct() %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct(),  
                         goAnnotations.sub2 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(goTerm) %>% distinct()
                         ) %>% nrow(), 
  category = c("Maize1", "Maize2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=GOTerms_by_subgenome_a), top="GO Term by Subgenome\n(EV-EXP, MF only)")

## Only keep genes where there is a duplicate still on subgenome 1 and subgenome 2.  The converted data.frame will hold maize2 genes with the names replaced with maize1 homologs
converted <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  right_join(., goAnnotations.sub2,by=c("Maize2"="gene2"))
converted$Maize2[!is.na(converted$Maize1)] <- as.character(converted$Maize1[!is.na(converted$Maize1)])
converted <- 
  converted %>%
  select(Maize2, goTerm, evCode, type) %>%
  rename(gene2=Maize2)

#EXP only, MF only
grid.newpage()
GOTerms_by_subgenome_b <- draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% nrow(),
  area2 = converted %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(gene2, goTerm) %>% distinct(), 
                         converted %>% filter(evCode=="EXP") %>% filter(type=="MF") %>% select(gene2, goTerm) %>% distinct()
                         ) %>% nrow(), 
  category = c("Maize1", "Maize2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=GOTerms_by_subgenome_b), top="GO Term Assignments by Subgenome\n(EV-EXP, MF only)")

```

# 3 Isoforms

***

Here is a scatter plot of isoform counts across gene pairs.

```{r subgenome_isoforms_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.align='center'}
# Isoforms
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(geneTranscript.counts, by=c("Maize1"="gene")) %>%
  inner_join(geneTranscript.counts, by=c("Maize2"="gene"))
names(data)[3] <- "Value_maize1"
names(data)[4] <- "Value_maize2"
# data <- subset(data, data$Value_maize1 > 1 & data$Value_maize2 > 1) ## Ignore single-exon genes
data$foldChange <- log2(data$Value_maize1) - log2(data$Value_maize2)

n <-
  data %>%
  # subset(Value_maize1 > 100 | Value_maize2 > 100) %>%
  subset(Value_maize1 > 25 | Value_maize2 > 25) %>%
  nrow()

subgenome_isoforms_scatter <-
  ggplot(subset(data, Value_maize1 <= 25 & Value_maize2 <= 25), aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(position = "jitter", alpha=.1, aes(color = abs(foldChange))) +
  scale_color_gradient(low = "blue", high = "red") +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  # xlim(0,100) +
  # ylim(0,100) +
  xlim(0,25) +
  ylim(0,25) +
  labs(
    # title = paste0("Scatterplot of isoform counts for gene pair(s).\nExcludes ",n," outlier, alpha=.1, foldchange in log2, blue=lm fit."),
    x = "Isoform Counts for Maize1 Gene",
    y = "Isoform Counts for Maize2 Gene"
  ) + 
  # annotate("text",x=85,y=20, label=paste("Pearson Cor = ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
  annotate("text",x=20,y=3, label=paste("Pearson Cor = ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
plot(subgenome_isoforms_scatter)

```

# 4 Expression

***

```{r subgenome_expression_scatter_rep, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=2.5, fig.align='center'}
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

data <- 
  data %>%
  subset(sample %in% c("Primary Root 5 Days", "Mature Leaf 8", "6-7 internode", "B73 Mature Pollen"))

subgenome_expression_scatter_rep <-
  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(alpha=.1, aes(color = sample)) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  scale_x_log10() + scale_y_log10() +
  labs(
    # title = paste0("Scatterplot of expression for gene pair(s).\nAlpha=.1, reference line"),
    x = "Expression for Maize1 Gene (log scale)",
    y = "Expression for Maize2 Gene (log scale)"
  ) +
  theme(legend.position = "none") + facet_wrap(~sample, nrow=1)
plot(subgenome_expression_scatter_rep)

```

```{r subgenome_expression_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

subgenome_expression_scatter <-
  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(alpha=.1, aes(color = sample)) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  scale_x_log10() + scale_y_log10() +
  labs(
    # title = paste0("Scatterplot of expression for gene pair(s).\nAlpha=.1, reference line"),
    x = "Expression for Maize1 Gene (log scale)",
    y = "Expression for Maize2 Gene (log scale)"
  ) +
  theme(legend.position = "none") + facet_wrap(~sample, nrow=6)
plot(subgenome_expression_scatter)
```

# 5 Protein Abundance

***

```{r subgenome_protein_scatter_rep, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=2.5, fig.align='center'}
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))
data <- 
  data %>%
  subset(sample %in% c("Primary Root 5 Days", "Mature Leaf 8", "6-7 internode", "B73 Mature Pollen"))

subgenome_protein_scatter_rep <-
  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(alpha=.1, aes(color = sample)) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  scale_x_log10() + scale_y_log10() +
  labs(
    # title = paste0("Scatterplot of protein abundance for gene pair(s).\nAlpha=.1, reference line"),
    x = "Abundance for Maize1 Gene (log scale)",
    y = "Abundance for Maize2 Gene (log scale)"
  ) +
  theme(legend.position = "none") + facet_wrap(~sample, nrow=1)
  # + annotate("text", size=2, x=10000,y=10, label=paste("pearson: ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
plot(subgenome_protein_scatter_rep)
```

```{r subgenome_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

subgenome_protein_scatter <-
  ggplot(data, aes(x=Value_maize1, y=Value_maize2)) +
  geom_point(alpha=.1, aes(color = sample)) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  scale_x_log10() + scale_y_log10() +
  labs(
    # title = paste0("Scatterplot of protein abundance for gene pair(s).\nAlpha=.1, reference line"),
    x = "Abundance for Maize1 Gene (log scale)",
    y = "Abundance for Maize2 Gene (log scale)"
  ) +
  theme(legend.position = "none") + facet_wrap(~sample, nrow=6)
  # + annotate("text", size=2, x=10000,y=10, label=paste("pearson: ", round(cor(data$Value_maize1, data$Value_maize2, method = c("pearson")),2)))
plot(subgenome_protein_scatter)
```

# 6 Expression and Abundance are Weakly Correlated

***

```{r subgenome_expression_protein_foldchange_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=6, fig.height=5, fig.align='center'}
## Start with the paired expression data
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "FPKM_maize1"
names(data)[5] <- "FPKM_maize2"

## Add paired abundance data
data <-
  data %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID", "sample"="sample")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[6] <- "dNSAF_maize1"
names(data)[7] <- "dNSAF_maize2"

# Remove pollen
data <-
  data %>%
  subset(!sample %in% c("B73 Mature Pollen"))

data$foldChange_expr <- log2(data$FPKM_maize1) - log2(data$FPKM_maize2)
data$foldChange_abun <- log2(data$dNSAF_maize1) - log2(data$dNSAF_maize2)
subgenome_expression_protein_foldchange_scatter <- ggplot(data, aes(foldChange_expr, foldChange_abun)) +
  geom_point(alpha=.1) +
  xlim(-10,10) + ylim(-10,10) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  # facet_wrap(~sample, nrow=6) +
  labs(
    # title = paste0("Scatterplot of Foldchange in Expression in Maize1 vs.\nFoldchange in Abundance in Maize1\nAlpha=.1, linear fit, n=13413"),
    x = "log2(Expression Maize1) - log2(Expression Maize2)",
    y = "log2(Abundance Maize1) - log2(Abundance Maize2)"
  ) + 
  annotate("text",x=7,y=-8, label=paste("Pearson Cor = ", round(cor(data$foldChange_expr, data$foldChange_abun, method = c("pearson")),2))) + 
  annotate("text",x=7,y=-10, label=paste("Spearman Cor = ", round(cor(data$foldChange_expr, data$foldChange_abun, method = c("spearman")),2)))
plot(subgenome_expression_protein_foldchange_scatter)
```

```{r subgenome_expression_protein_scatter_rep, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=2.5, fig.align='center'}
## Start with the paired expression data
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "FPKM_maize1"
names(data)[5] <- "FPKM_maize2"

## Add paired abundance data
data <-
  data %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID", "sample"="sample")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[6] <- "dNSAF_maize1"
names(data)[7] <- "dNSAF_maize2"

# Pollen behaves much differently than the rest, lets see
data <- 
  data %>%
  subset(sample %in% c("Primary Root 5 Days", "Mature Leaf 8", "6-7 internode", "B73 Mature Pollen"))

## View as scatter
subgenome_expression_protein_scatter_rep <- ggplot(data, aes(FPKM_maize1, dNSAF_maize1)) +
  geom_point(alpha=.1, aes(color = sample)) +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  # geom_smooth(method="loess", se=TRUE, fullrange=FALSE, level=0.95) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  scale_x_log10() + scale_y_log10() +
  theme(legend.position = "none") + facet_wrap(~sample, nrow=1) +
  labs(
    # title = paste0("Scatterplot of Expression in Maize1 vs. Abundance in Maize1\nLog-Log Scale for gene pair(s).\nAlpha=.1, reference line (black), lm fit (blue)"),
    x = "Expression for Maize1 Gene (log scale)",
    y = "Abundance for Maize1 Gene (log scale)"
  ) 
plot(subgenome_expression_protein_scatter_rep)
```

```{r subgenome_expression_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
## Start with the paired expression data
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "FPKM_maize1"
names(data)[5] <- "FPKM_maize2"

## Add paired abundance data
data <-
  data %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID", "sample"="sample")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[6] <- "dNSAF_maize1"
names(data)[7] <- "dNSAF_maize2"

## View as scatter
subgenome_expression_protein_scatter <- ggplot(data, aes(FPKM_maize1, dNSAF_maize1)) +
  geom_point(alpha=.1, aes(color = sample)) +
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95) +
  # geom_smooth(method="loess", se=TRUE, fullrange=FALSE, level=0.95) +
  geom_abline(mapping = null, data = null, slope = 1, intercept = 0) +
  scale_x_log10() + scale_y_log10() +
  theme(legend.position = "none") + facet_wrap(~sample, nrow=6) +
  labs(
    # title = paste0("Scatterplot of Expression in Maize1 vs. Abundance in Maize1\nLog-Log Scale for gene pair(s).\nAlpha=.1, reference line (black), lm fit (blue)"),
    x = "Expression for Maize1 Gene (log scale)",
    y = "Abundance for Maize1 Gene (log scale)"
  ) 
plot(subgenome_expression_protein_scatter)
```

# 7 Expression pattern types
functionalDivergence.R

# 8 Misc

***

```{r extra1, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=4, fig.height=4, fig.align='center'}
#Pollen has far fewer cases with exp
## Start with the paired expression data
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "FPKM_maize1"
names(data)[5] <- "FPKM_maize2"

case_counts <- data %>% group_by(sample) %>% count() %>% arrange(n)
ggplot(case_counts, aes(reorder(sample, n), n)) + geom_col() + theme(axis.text.x=element_text(angle=90,hjust=1))
```

```{r extra2, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=4, fig.height=4, fig.align='center'}
#Pollen has far fewer cases with abun
## Start with the abundance data
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[4] <- "dNSAF_maize1"
names(data)[5] <- "dNSAF_maize2"

case_counts <- data %>% group_by(sample) %>% count() %>% arrange(n)
ggplot(case_counts, aes(reorder(sample, n), n)) + geom_col() + theme(axis.text.x=element_text(angle=90,hjust=1))
```

```{r extra3, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=4, fig.height=4, fig.align='center'}
#Pollen has far fewer cases with both exp and abun
## Start with the paired expression data
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "FPKM_maize1"
names(data)[5] <- "FPKM_maize2"

## Add paired abundance data
data <-
  data %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID", "sample"="sample")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[6] <- "dNSAF_maize1"
names(data)[7] <- "dNSAF_maize2"

case_counts <- data %>% group_by(sample) %>% count() %>% arrange(n)
ggplot(case_counts, aes(reorder(sample, n), n)) + geom_col() + theme(axis.text.x=element_text(angle=90,hjust=1))
```

***

# 9 Wrapup

***

```{r print, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center'}
ggsave("Output/Figure 1.png", plot = chromosome_plot)

png("Output/Figure 2a.png");
grid.draw(GOTerms_by_subgenome_a)
dev.off()
png("Output/Figure 2b.png");
grid.draw(GOTerms_by_subgenome_b)
dev.off()

ggsave("Output/Figure 3.png", plot = subgenome_isoforms_scatter)
ggsave("Output/Figure 4.png", plot = subgenome_expression_scatter_rep)
ggsave("Output/Figure 5.png", plot = subgenome_protein_scatter_rep)
ggsave("Output/Figure 6.png", plot = subgenome_expression_protein_foldchange_scatter)
ggsave("Output/Figure 7.png", plot = subgenome_expression_protein_scatter)
ggsave("Output/Figure 8.png", plot = subgenome_expression_protein_scatter_rep)
ggsave("Output/Figure S1.png", plot = subgenome_expression_scatter)
ggsave("Output/Figure S2.png", plot = subgenome_protein_scatter)
ggsave("Output/Figure S3.png", plot = subgenome_expression_protein_scatter)
```

Report generated on: `r as.character(format(Sys.Date(), format="%B %d, %Y"))`